---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Development of this package has been abandoned

The main function in this package has now been incorporated into the official 
version of the 
[htmlTable](https://cran.r-project.org/web/packages/htmlTable/index.html)
package, and development of this package will not longer be supported.

In the `htmlTable` package, the `htmlTable_td` function has been renamed 
to `tidyHtmlTable` with  additional changes made to parameter names.

See `?tidyHtmlTable` for further details.

```{r gh-installation, eval=FALSE}

# For the dev version of htmlTable
install.packages("devtools")
devtools::install_github("gforge/tidyhTbl")

# For the lastest CRAN release
install.packages("htmlTable")

```


[![Build Status](https://travis-ci.org/graggsd/tidyhTbl.svg?branch=master)](https://travis-ci.org/graggsd/tidyhTbl)

# tidyhTbl

The goal of tidyhTbl is to wrap the `htmlTable` function from the 
[htmlTable](https://cran.r-project.org/web/packages/htmlTable/index.html)
package for use with tidy data.

## Installation

You can install tidyhTbl from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("graggsd/tidyhTbl")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, message=FALSE}
library(magrittr)
library(tidyr)
library(dplyr)
library(tidyhTbl)
library(tibble)
mtcars %>%
    rownames_to_column %>%
    select(rowname, cyl, gear, hp, mpg, qsec) %>%
    gather(per_metric, value, hp, mpg, qsec) %>%
    group_by(cyl, gear, per_metric) %>%
    summarise(Mean = round(mean(value), 1),
              SD = round(sd(value), 1),
              Min = round(min(value), 1),
              Max = round(max(value), 1)) %>%
     gather(summary_stat, value, Mean, SD, Min, Max) %>%
     ungroup %>%
     mutate(gear = paste(gear, "Gears"),
            cyl = paste(cyl, "Cylinders")) %>%
     htmlTable_td(header_td = "gear",
                  cgroup1_td = "cyl",
                  cell_value = "value",
                  rnames_td = "summary_stat",
                  rgroup_td = "per_metric")
```
